---
title: "Reanalysis for Problem Set 2"
author: "Erin Bennett"
date: "February 16, 2016"
output: 
  html_document:
      toc: true
---

```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, cache=TRUE, message=FALSE, sanitiz =TRUE)
```

**Caveat:** I probably made mistakes in this code...

**Fact:** [Sklar et al. (2012)](http://www.pnas.org/content/109/48/19614.full) find that for their subtraction problems, reaction time is lower for primed targets than unprimed targets.

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
library(lme4)
library(ggthemes)

sem <- function(x) {sd(x, na.rm=TRUE) / sqrt(length(x))}

# for bootstrapping 95% confidence intervals
theta <- function(x,xdata) {mean(xdata[x])}
ci.low <- function(x) {
  quantile(bootstrap::bootstrap(1:length(x),1000,theta,x)$thetastar,.025)}
ci.high <- function(x) {
  quantile(bootstrap::bootstrap(1:length(x),1000,theta,x)$thetastar,.975)}

theme_new <- theme_set(theme_few())
```

# Experiment 6

```{r echo=F, message=F, warning=F}
subinfo <- read.csv("http://langcog.stanford.edu/sklar_expt6_subinfo_corrected.csv")
d.a <- read.csv("http://langcog.stanford.edu/sklar_expt6a_corrected.csv")
d.b <- read.csv("http://langcog.stanford.edu/sklar_expt6b_corrected.csv")

subj.id.columns.a = which(names(d.a) %in% paste('X', 1:21, sep=''))
subj.id.columns.b = which(names(d.b) %in% paste('X', 22:42, sep=''))
d.a$trial.order = 1
d.b$trial.order = 2
d.both = rbind(
  d.a %>% gather('subid', 'rt', subj.id.columns.a),
  d.b %>% gather('subid', 'rt', subj.id.columns.b) ) %>%
  mutate(subid = factor(gsub('X', '', as.character(subid))),
         trial.order = factor(trial.order))

d6 = merge(d.both, subinfo, 'subid')

d6$presentation.time <- factor(d6$presentation.time)
levels(d6$operand) <- c("addition","subtraction")

d6 = filter(d6, !is.na(rt))
```

## Design

For each prime, there were as many congruent targets as there were incongruent targets (though when we filter out datapoints where the reaction time was actually recorded, we get slight asymmetries). However, the result of the primes were different for the different operations, and the subtraction results were skewed to be close to (and frequently equal to) zero.

```{r echo=F, message=F, warning=F, fig.width=7, fig.height=3}
ggplot(d6, aes(x=prime.result, fill=congruent)) +
  geom_bar(stat='bin', position='dodge', binwidth=0.5) +
  facet_wrap(~operand) +
  ggtitle('frequency of results of the priming equations with measured rt')
```

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
d6 %>% group_by(operand, prime.result, target) %>%
  summarise(N = length(prime.result)) %>%
  ggplot(., aes(x=prime.result, y=target, fill=N)) +
  geom_tile() +
  facet_wrap(~operand, scale='free') +
  scale_fill_gradient(low = 'navy', high = 'red') +
  ggtitle('frequency of pairings of prime results with targets')
```

## Problems with zero trials?

**Hypothesis based on design:** Maybe the effect is driven by trials where the prime.result is 0?

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
ds6 = d6 %>% filter(subjective.test == 0 & objective.test <= 0.6)
ds6 %>%
  mutate(prime.result.is.0 = prime.result==0) %>%
  group_by(prime.result.is.0, operand, presentation.time, subid, congruent) %>%
  summarise(rt = mean(rt, na.rm=T)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(sem = sem(difference.score),
            difference.score = mean(difference.score)) %>%
  ggplot(., aes(x=presentation.time, y=difference.score,
                colour=prime.result.is.0,
                fill=prime.result.is.0)) +
  geom_bar(stat='identity', position='dodge', width=0.3, alpha=1/3) +
  geom_errorbar(aes(ymin=difference.score-sem, ymax=difference.score+sem), width=0.05) +
  geom_hline(yintercept=0, colour='gray') +
  facet_wrap(operand ~ prime.result.is.0)
```

```{r echo=F, message=F, warning=F, fig.width=5, fig.height=3}
by.quantity.6.excl <- ds6 %>%
  group_by(prime.result, operand, prime, congruent) %>%
  summarise(rt = mean(rt), 
            n = n()) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes'],
            n = sum(n)) %>%
  summarise(mean.ds = mean(difference.score),
            cil = ci.low(difference.score), 
            cih = ci.high(difference.score), 
            n = sum(n)) 

ggplot(by.quantity.6.excl, aes(x=prime.result, y=mean.ds)) +
  ylab('mean difference score') +
  xlab('primed result') +
  geom_hline(yintercept=0) +
  geom_errorbar(aes(x=prime.result, ymin=cil, ymax=cih), width=0) +
  geom_point(pch=21, aes(size = n), fill='white') +
  geom_text(aes(label=prime.result), size=3)
```

The results look similar whether or not we use Sklar et al's exclusion criteria: unfiltered.

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
d6 %>%
  mutate(prime.result.is.0 = prime.result==0) %>%
  group_by(prime.result.is.0, operand, presentation.time, subid, congruent) %>%
  summarise(rt = mean(rt, na.rm=T)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(sem = sem(difference.score),
            difference.score = mean(difference.score)) %>%
  ggplot(., aes(x=presentation.time, y=difference.score,
                colour=prime.result.is.0,
                fill=prime.result.is.0)) +
  geom_bar(stat='identity', position='dodge', width=0.3, alpha=1/3) +
  geom_errorbar(aes(ymin=difference.score-sem, ymax=difference.score+sem), width=0.05) +
  geom_hline(yintercept=0, colour='gray') +
  facet_wrap(operand ~ prime.result.is.0)
```

```{r echo=F, message=F, warning=F, fig.width=5, fig.height=3}
by.quantity.6.all <- d6 %>%
  group_by(prime.result, operand, prime, congruent) %>%
  summarise(rt = mean(rt), 
            n = n()) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes'],
            n = sum(n)) %>%
  summarise(mean.ds = mean(difference.score),
            cil = ci.low(difference.score), 
            cih = ci.high(difference.score), 
            n = sum(n)) 

ggplot(by.quantity.6.all, aes(x=prime.result, y=mean.ds)) +
  ylab('mean difference score') +
  xlab('primed result') +
  geom_hline(yintercept=0) +
  geom_errorbar(aes(x=prime.result, ymin=cil, ymax=cih), width=0) +
  geom_point(pch=21, aes(size = n), fill='white') +
  geom_text(aes(label=prime.result), size=3)
```

## Easiness of prime?

**Hypothesis about mechanism:** Does the "easiness" of the prime predict to the fascilitation effect?

Well, the other highest difference scores are when the answer is 10, which also seems easy...

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
d6$prime.result.class='other'
d6$prime.result.class[d6$prime.result==0] = '0'
d6$prime.result.class[d6$prime.result==10] = '10'
ds6 = d6 %>% filter(subjective.test == 0 & objective.test <= 0.6)
ds6 %>%
  group_by(prime.result.class, operand, presentation.time, subid, congruent) %>%
  summarise(rt = mean(rt, na.rm=T)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(sem = sem(difference.score),
            difference.score = mean(difference.score)) %>%
  ggplot(., aes(x=presentation.time, y=difference.score,
                colour=prime.result.class,
                fill=prime.result.class)) +
  geom_bar(stat='identity', position='dodge', width=0.3, alpha=1/3) +
  geom_errorbar(aes(ymin=difference.score-sem, ymax=difference.score+sem), width=0.05) +
  geom_hline(yintercept=0, colour='gray') +
  facet_wrap(operand ~ prime.result.class)
```

We could go on mturk and get the time it takes to solve each of these priming equations. Does that easiness correlate with the difference score?

# Experiment 7

```{r echo=F, message=F, warning=F}
subinfo <- read.csv("sklar_expt7_subinfo.csv")
d.a <- read.csv("sklar_expt7a.csv")
d.b <- read.csv("sklar_expt7b.csv")

subj.id.columns.a = which(names(d.a) %in% paste('X', 1:100, sep=''))
subj.id.columns.b = which(names(d.b) %in% paste('X', 1:100, sep=''))
d.a$trial.order = 1
d.b$trial.order = 2
d.both = rbind(
  d.a %>% gather('subid', 'rt', subj.id.columns.a),
  d.b %>% gather('subid', 'rt', subj.id.columns.b) ) %>%
  mutate(subid = factor(gsub('X', '', as.character(subid))),
         trial.order = factor(trial.order))

d7 = merge(d.both, subinfo, 'subid')

d7$presentation.time <- factor(d7$presentation.time)
levels(d7$operand) <- c("addition","subtraction")

d7 = filter(d7, !is.na(rt)) %>%
  filter(prime != '7+2+3=' & prime != '3+8+1=')
```

## Design

The design of experiment 7 was much more balanced.

```{r echo=F, message=F, warning=F, fig.width=7, fig.height=3}
ggplot(d7, aes(x=prime.result, fill=congruent)) +
  geom_histogram(stat='bin', position='dodge', binwidth=0.5) +
  facet_wrap(~operand) +
  ggtitle('frequency of results of the priming equations with measured rt')
```

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
d7 %>% group_by(operand, prime.result, target) %>%
  summarise(N = length(prime.result)) %>%
  ggplot(., aes(x=prime.result, y=target, fill=N)) +
  geom_tile() +
  facet_wrap(~operand, scale='free') +
  scale_fill_gradient(low = 'navy', high = 'red') +
  ggtitle('frequency of pairings of prime results with targets')
```

## Effects

When we don't exlude we see the effect of 0.

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
d7 %>%
  mutate(prime.result.is.0 = prime.result==0) %>%
  group_by(prime.result.is.0, operand, presentation.time, subid, congruent) %>%
  summarise(rt = mean(rt, na.rm=T)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(sem = sem(difference.score),
            difference.score = mean(difference.score)) %>%
  ggplot(., aes(x=presentation.time, y=difference.score,
                colour=prime.result.is.0,
                fill=prime.result.is.0)) +
  geom_bar(stat='identity', position='dodge', width = 1/3, alpha=1/3) +
  geom_errorbar(aes(ymin=difference.score-sem, 
                    ymax=difference.score+sem), width=0.05) +
  geom_hline(yintercept=0, colour='gray') +
  facet_wrap(operand ~ prime.result.is.0)
```

```{r echo=F, message=F, warning=F, fig.width=5, fig.height=3}
by.quantity.7.all <- d7 %>%
  group_by(prime.result, operand, prime, congruent) %>%
  summarise(rt = mean(rt), 
            n = n()) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes'],
            n = sum(n)) %>%
  summarise(mean.ds = mean(difference.score),
            cil = ci.low(difference.score), 
            cih = ci.high(difference.score), 
            n = sum(n)) 

ggplot(by.quantity.7.all, aes(x=prime.result, y=mean.ds)) +
  ylab('mean difference score') +
  xlab('primed result') +
  geom_hline(yintercept=0) +
  geom_errorbar(aes(x=prime.result, ymin=cil, ymax=cih), width=0) +
  geom_point(pch=21, aes(size = n), fill='white') +
  geom_text(aes(label=prime.result), size=3)
```

Same when we do exclude, but less strong. 

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
d7$prime.result.class='other'
d7$prime.result.class[d7$prime.result==0] = '0'
d7$prime.result.class[d7$prime.result==10] = '10'
ds7 = d7 %>% filter(subjective.test==1 & objective.test<0.6)

ds7 %>%
  mutate(prime.result.is.0 = prime.result==0) %>%
  group_by(prime.result.is.0, operand, presentation.time, subid, congruent) %>%
  summarise(rt = mean(rt, na.rm=T)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(sem = sem(difference.score),
            difference.score = mean(difference.score)) %>%
  ggplot(., aes(x=presentation.time, y=difference.score,
                colour=prime.result.is.0,
                fill=prime.result.is.0)) +
  geom_bar(stat='identity', position='dodge', width=0.3, alpha=1/3) +
  geom_errorbar(aes(ymin=difference.score-sem, ymax=difference.score+sem), width=0.05) +
  geom_hline(yintercept=0, colour='gray') +
  facet_wrap(operand ~ prime.result.is.0)
```

```{r echo=F, message=F, warning=F, fig.width=5, fig.height=3}
by.quantity.7.excl <- ds7 %>%
  group_by(prime.result, operand, prime, congruent) %>%
  summarise(rt = mean(rt), 
            n = n()) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes'],
            n = sum(n)) %>%
  summarise(mean.ds = mean(difference.score),
            cil = ci.low(difference.score), 
            cih = ci.high(difference.score), 
            n = sum(n)) 

ggplot(by.quantity.7.excl, aes(x=prime.result, y=mean.ds)) +
  ylab('mean difference score') +
  xlab('primed result') +
  geom_hline(yintercept=0) +
  geom_errorbar(aes(x=prime.result, ymin=cil, ymax=cih), width=0) +
  geom_point(pch=21, aes(size = n), fill='white') +
  geom_text(aes(label=prime.result), size=3)
```

10 has a lower fasciliation effect than in experiment 6.

# Cross-experiments

```{r}
ms <- bind_rows(by.quantity.6.excl %>% 
                  mutate(expt = "e6", 
                         excl = TRUE),
                by.quantity.6.all %>%
                  mutate(expt = "e6", 
                         excl = FALSE),
                by.quantity.7.excl %>%
                  mutate(expt = "e7", 
                         excl = TRUE),
                by.quantity.7.all %>% 
                  mutate(expt = "e7", 
                         excl = FALSE)) 
  
ggplot(ms, aes(x=prime.result, y=mean.ds)) +
  ylab('mean difference score') +
  xlab('primed result') +
  geom_hline(yintercept=0) +
  geom_errorbar(aes(x=prime.result, ymin=cil, ymax=cih), width=0) +
  geom_point(pch=21, aes(size = n), fill='white') +
  geom_text(aes(label=prime.result), size=3) + 
  facet_grid(expt ~ excl)
```

Not much correlation between experiments in individual values. With exclusions:

```{r}
ms %>% 
  filter(excl) %>%
  select(-cil, -cih, -n) %>%
  spread(expt, mean.ds) %>%
  ggplot(aes(x = e6, y = e7, col = operand, label = prime.result)) +
  geom_text() + 
  geom_smooth(method = "lm")
```

Without.

```{r}
ms %>% 
  filter(!excl) %>%
  select(-cil, -cih, -n) %>%
  spread(expt, mean.ds) %>%
  ggplot(aes(x = e6, y = e7, col = operand, label = prime.result)) +
  geom_text() + 
  geom_smooth(method = "lm")
```