---
title: "Reanalysis for Problem Set 2"
author: "Erin Bennett"
date: "February 16, 2016"
output: html_document
---

# Intro

**Caveat:** I probably made mistakes in this code...

**Fact:** [Sklar et al. (2012)](http://www.pnas.org/content/109/48/19614.full) find that for their subtraction problems, reaction time is lower for primed targets than unprimed targets.

```{r echo=F, message=F, warning=F}
library(ggplot2)
library(tidyr)
library(dplyr)
library(lme4)
library(ggthemes)
sem <- function(x) {sd(x, na.rm=TRUE) / sqrt(length(x))}
ci95 <- function(x) {sem(x) * 1.96}

theme_new <- theme_set(theme_few())
char = as.character
num = function(x) {return(as.numeric(as.character(x)))}

subinfo <- read.csv("http://langcog.stanford.edu/sklar_expt6_subinfo_corrected.csv")
d.a <- read.csv("http://langcog.stanford.edu/sklar_expt6a_corrected.csv")
d.b <- read.csv("http://langcog.stanford.edu/sklar_expt6b_corrected.csv")

subj.id.columns.a = which(names(d.a) %in% paste('X', 1:21, sep=''))
subj.id.columns.b = which(names(d.b) %in% paste('X', 22:42, sep=''))
d.a$trial.order = 1
d.b$trial.order = 2
d.both = rbind(
  d.a %>% gather('subid', 'rt', subj.id.columns.a),
  d.b %>% gather('subid', 'rt', subj.id.columns.b) ) %>%
  mutate(subid = factor(gsub('X', '', as.character(subid))),
         trial.order = factor(trial.order))

d = merge(d.both, subinfo, 'subid')

d$presentation.time <- factor(d$presentation.time)
levels(d$operand) <- c("addition","subtraction")

d = filter(d, !is.na(rt))
```

# Design

For each prime, there were as many congruent targets as there were incongruent targets (though when we filter out datapoints where the reaction time was actually recorded, we get slight asymmetries). However, the result of the primes were different for the different operations, and the subtraction results were skewed to be close to (and frequently equal to) zero.

```{r echo=F, message=F, warning=F, fig.width=7, fig.height=3}
ggplot(d, aes(x=prime.result, fill=congruent)) +
  geom_bar(stat='bin', position='dodge', binwidth=0.5) +
  facet_wrap(~operand) +
  ggtitle('frequency of results of the priming equations with measured rt')
```

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
d %>% group_by(operand, prime.result, target) %>%
  summarise(N = length(prime.result)) %>%
  ggplot(., aes(x=prime.result, y=target, fill=N)) +
  geom_tile() +
  facet_wrap(~operand, scale='free') +
  scale_fill_gradient(low = 'navy', high = 'red') +
  ggtitle('frequency of pairings of prime results with targets')
```

**Hypothesis based on design:** Maybe the effect is driven by trials where the prime.result is 0?

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
ds = d %>% filter(subjective.test == 0 & objective.test <= 0.6)
ds %>%
  mutate(prime.result.is.0 = prime.result==0) %>%
  group_by(prime.result.is.0, operand, presentation.time, subid, congruent) %>%
  summarise(rt = mean(rt, na.rm=T)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(sem = sem(difference.score),
            difference.score = mean(difference.score)) %>%
  ggplot(., aes(x=presentation.time, y=difference.score,
                colour=prime.result.is.0,
                fill=prime.result.is.0)) +
  geom_bar(stat='identity', position='dodge', width=0.3, alpha=1/3) +
  geom_errorbar(aes(ymin=difference.score-sem, ymax=difference.score+sem), width=0.05) +
  geom_hline(y=0, colour='gray') +
  facet_wrap(operand ~ prime.result.is.0)
```

```{r echo=F, message=F, warning=F, fig.width=5, fig.height=3}
ds %>%
  group_by(prime.result, operand, prime, congruent) %>%
  summarise(rt = mean(rt)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(mean.ds = mean(difference.score),
            sem.ds = sem(difference.score)) %>%
  ggplot(., aes(x=prime.result, y=mean.ds)) +
  ylab('mean difference score') +
  xlab('primed result') +
  geom_hline(y=0) +
  geom_errorbar(aes(x=prime.result, ymin=mean.ds-sem.ds, ymax=mean.ds+sem.ds), width=0) +
  geom_point(pch=21, size=6, fill='white') +
  geom_text(aes(label=prime.result), size=3)
```

The results look similar whether or not we use Sklar et al's exclusion criteria.

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
d %>%
  mutate(prime.result.is.0 = prime.result==0) %>%
  group_by(prime.result.is.0, operand, presentation.time, subid, congruent) %>%
  summarise(rt = mean(rt, na.rm=T)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(sem = sem(difference.score),
            difference.score = mean(difference.score)) %>%
  ggplot(., aes(x=presentation.time, y=difference.score,
                colour=prime.result.is.0,
                fill=prime.result.is.0)) +
  geom_bar(stat='identity', position='dodge', width=0.3, alpha=1/3) +
  geom_errorbar(aes(ymin=difference.score-sem, ymax=difference.score+sem), width=0.05) +
  geom_hline(y=0, colour='gray') +
  facet_wrap(operand ~ prime.result.is.0)
```

```{r echo=F, message=F, warning=F, fig.width=5, fig.height=3}
d %>%
  group_by(prime.result, operand, prime, congruent) %>%
  summarise(rt = mean(rt)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(mean.ds = mean(difference.score),
            sem.ds = sem(difference.score)) %>%
  ggplot(., aes(x=prime.result, y=mean.ds)) +
  ylab('mean difference score') +
  xlab('primed result') +
  geom_hline(y=0) +
  geom_errorbar(aes(x=prime.result, ymin=mean.ds-sem.ds, ymax=mean.ds+sem.ds), width=0) +
  geom_point(pch=21, size=6, fill='white') +
  geom_text(aes(label=prime.result), size=3)
```

**Hypothesis about mechanism:** Does the "easiness" of the prime predict to the fascilitation effect?

Well, the other highest difference scores are when the answer is 10, which also seems easy...

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
d$prime.result.class='other'
d$prime.result.class[d$prime.result==0] = '0'
d$prime.result.class[d$prime.result==10] = '10'
ds = d %>% filter(subjective.test == 0 & objective.test <= 0.6)
ds %>%
  group_by(prime.result.class, operand, presentation.time, subid, congruent) %>%
  summarise(rt = mean(rt, na.rm=T)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(sem = sem(difference.score),
            difference.score = mean(difference.score)) %>%
  ggplot(., aes(x=presentation.time, y=difference.score,
                colour=prime.result.class,
                fill=prime.result.class)) +
  geom_bar(stat='identity', position='dodge', width=0.3, alpha=1/3) +
  geom_errorbar(aes(ymin=difference.score-sem, ymax=difference.score+sem), width=0.05) +
  geom_hline(y=0, colour='gray') +
  facet_wrap(operand ~ prime.result.class)
```

We could go on mturk and get the time it takes to solve each of these priming equations. Does that easiness correlate with the difference score?

# Experiment 7

```{r echo=F, message=F, warning=F}
subinfo <- read.csv("sklar_expt7_subinfo.csv")
d.a <- read.csv("sklar_expt7a.csv")
d.b <- read.csv("sklar_expt7b.csv")

subj.id.columns.a = which(names(d.a) %in% paste('X', 1:100, sep=''))
subj.id.columns.b = which(names(d.b) %in% paste('X', 1:100, sep=''))
d.a$trial.order = 1
d.b$trial.order = 2
d.both = rbind(
  d.a %>% gather('subid', 'rt', subj.id.columns.a),
  d.b %>% gather('subid', 'rt', subj.id.columns.b) ) %>%
  mutate(subid = factor(gsub('X', '', as.character(subid))),
         trial.order = factor(trial.order))

d = merge(d.both, subinfo, 'subid')

d$presentation.time <- factor(d$presentation.time)
levels(d$operand) <- c("addition","subtraction")

d = filter(d, !is.na(rt)) %>%
  filter(prime != '7+2+3=' & prime != '3+8+1=')
```

The design of experiment 7 was much more balanced.

```{r echo=F, message=F, warning=F, fig.width=7, fig.height=3}
ggplot(d, aes(x=prime.result, fill=congruent)) +
  geom_bar(stat='bin', position='dodge', binwidth=0.5) +
  facet_wrap(~operand) +
  ggtitle('frequency of results of the priming equations with measured rt')
```

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
d %>% group_by(operand, prime.result, target) %>%
  summarise(N = length(prime.result)) %>%
  ggplot(., aes(x=prime.result, y=target, fill=N)) +
  geom_tile() +
  facet_wrap(~operand, scale='free') +
  scale_fill_gradient(low = 'navy', high = 'red') +
  ggtitle('frequency of pairings of prime results with targets')
```

When we don't exlude we see the effect of 0.

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
d %>%
  mutate(prime.result.is.0 = prime.result==0) %>%
  group_by(prime.result.is.0, operand, presentation.time, subid, congruent) %>%
  summarise(rt = mean(rt, na.rm=T)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(sem = sem(difference.score),
            difference.score = mean(difference.score)) %>%
  ggplot(., aes(x=presentation.time, y=difference.score,
                colour=prime.result.is.0,
                fill=prime.result.is.0)) +
  geom_bar(stat='identity', position='dodge', width=0.3, alpha=1/3) +
  geom_errorbar(aes(ymin=difference.score-sem, ymax=difference.score+sem), width=0.05) +
  geom_hline(y=0, colour='gray') +
  facet_wrap(operand ~ prime.result.is.0)
```

```{r echo=F, message=F, warning=F, fig.width=5, fig.height=3}
d %>%
  group_by(prime.result, operand, prime, congruent) %>%
  summarise(rt = mean(rt)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(mean.ds = mean(difference.score),
            sem.ds = sem(difference.score)) %>%
  ggplot(., aes(x=prime.result, y=mean.ds)) +
  ylab('mean difference score') +
  xlab('primed result') +
  geom_hline(y=0) +
  geom_errorbar(aes(x=prime.result, ymin=mean.ds-sem.ds, ymax=mean.ds+sem.ds), width=0) +
  geom_point(pch=21, size=6, fill='white') +
  geom_text(aes(label=prime.result), size=3)
```

Same when we do exclude, but less strong. 

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=3}
d$prime.result.class='other'
d$prime.result.class[d$prime.result==0] = '0'
d$prime.result.class[d$prime.result==10] = '10'
ds = d %>% filter(subjective.test==1 & objective.test<0.6)
ds %>%
  mutate(prime.result.is.0 = prime.result==0) %>%
  group_by(prime.result.is.0, operand, presentation.time, subid, congruent) %>%
  summarise(rt = mean(rt, na.rm=T)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(sem = sem(difference.score),
            difference.score = mean(difference.score)) %>%
  ggplot(., aes(x=presentation.time, y=difference.score,
                colour=prime.result.is.0,
                fill=prime.result.is.0)) +
  geom_bar(stat='identity', position='dodge', width=0.3, alpha=1/3) +
  geom_errorbar(aes(ymin=difference.score-sem, ymax=difference.score+sem), width=0.05) +
  geom_hline(y=0, colour='gray') +
  facet_wrap(operand ~ prime.result.is.0)
```

```{r echo=F, message=F, warning=F, fig.width=5, fig.height=3}
ds %>%
  group_by(prime.result, operand, prime, congruent) %>%
  summarise(rt = mean(rt)) %>%
  summarise(difference.score = rt[congruent=='no']-rt[congruent=='yes']) %>%
  summarise(mean.ds = mean(difference.score),
            sem.ds = sem(difference.score)) %>%
  ggplot(., aes(x=prime.result, y=mean.ds)) +
  ylab('mean difference score') +
  xlab('primed result') +
  geom_hline(y=0) +
  geom_errorbar(aes(x=prime.result, ymin=mean.ds-sem.ds, ymax=mean.ds+sem.ds), width=0) +
  geom_point(pch=21, size=6, fill='white') +
  geom_text(aes(label=prime.result), size=3)
```

10 has a lower fasciliation effect than in experiment 6.